---
title: "eTRF offspring body composition analysis"
author: "Molly C. Mulcahy"
date: "2/25/2021"
output:
  html_document: default
  pdf_document: default
---
#This document contains primary data and analysis for the dissertation work involving early time-restricted feeding during gestation effect on mouse offspring

#setup chunk
```{r setup, include=FALSE}
library(processx)
library(devtools)
library(dplyr)
library(readr)
library(car)
library(knitr)
library(tidyr)
library(lme4)
library(broom)
library(RCurl)
library(ggplot2)

#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively
knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})
#write standard error function for later use
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
# sets color scheme for plots
color.scheme <- c('gray69','gray36')
```

#Data Entry
```{r data-entry-MRI}
MRI.data.file<-"MRI.Off.data.csv"
MRI.raw.data<-read_csv(MRI.data.file,
                       col_types = cols(
  ID = col_factor(levels = NULL),
  age = col_double(),
  animal.id = col_factor(levels = NULL),
  MouseID = col_factor(levels = NULL),
  Sex = col_factor(levels = c("M","F")),
  Genotype = col_factor(levels = NULL),
  values = col_double(),
  assay = col_factor(levels = NULL),
  experiment.date = col_date(format = ""),
  treatment = col_factor(levels = c("eTRF","AL")),
  dam = col_factor(levels = NULL),
  cohort = col_factor(levels = NULL),
  gram.values = col_double()))

MRI.raw.data<-MRI.raw.data%>%
  mutate(sex = case_when(
    Sex=="F"~"female",
    Sex =="M"~"male"))

cage.file<-"cage_mapping_offspring.csv"
cage.data<-read_csv(cage.file,
                    col_types = cols(
                      ID = col_factor(levels = NULL),
                      treatment = col_factor(levels = c("eTRF", "AL")),
                      dam = col_factor(levels = NULL),
                      cohort = col_factor(levels = NULL),
                      cage = col_factor(levels = NULL)
                    ))

MRI.raw.data<-full_join(cage.data, MRI.raw.data, by = c("ID", "treatment","dam","cohort"))
```
#Pre-HFD analysis (PND 21- PND 70)
#Body Composition
```{r Figure2A}
#table of unique animals for body composition numbers
MRI.raw.data %>%
  group_by(sex,treatment) %>%
  distinct(ID, .keep_all = T) %>%
  count %>%
  kable(caption="Animals in each group of this cohort")

#Body Weight data table
early.HFD<-c("8673","8674","8688")
BW.data<-filter(MRI.raw.data, assay=="Body Weight")%>%
  filter(age<70)%>%
  filter(!(age>=67&ID%in%early.HFD))%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
##Body Weight plot
ggplot(BW.data,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Body Weight", y = "Body weight (grams)",x="Age (days)")+
  theme(text = element_text(size=18),legend.position = c(0.9,0.19), legend.background = element_rect(fill = "transparent"))

#all
BW.data.all<-MRI.raw.data%>%
  filter(assay=="Body Weight")%>%
   filter(age<70)%>%
  filter(!(age>=67&ID%in%early.HFD))

null.bw.lme<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = BW.data.all)
age.bw.lme<-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = BW.data.all)
anova(null.bw.lme, age.bw.lme)#significant effect of age
diet.bw.lme<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = BW.data.all)

sex.bw.lme<-lmer(gram.values~ age + treatment*sex +(1|MouseID) + (1|dam), data = BW.data.all) 
anova(sex.bw.lme, diet.bw.lme)#p<0.0001
Anova(sex.bw.lme)%>%tidy()%>%kable()
anova(age.bw.lme, diet.bw.lme)%>%tidy()%>%kable()#p=0.471
#anova(diet.bw.lme)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
```
There is a significant interaction of age and treatment in females where age-matched AL female have a ,p-value.

```{r Fig2C}
#Lean Mass data table
LM.data<-filter(MRI.raw.data, assay=="Lean Mass")%>%
  filter(age<70)%>%
  filter(!(age>=67&ID%in%early.HFD))%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
#Lean Mass Plot
ggplot(LM.data,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Lean Mass", y = "Lean Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=18),legend.position = c(0.88,0.20), legend.background = element_rect(fill = "transparent"))

#stats
LM.data.all<-MRI.raw.data%>%
  filter(assay=="Lean Mass")%>%
    filter(age<70)%>%
  filter(!(age>=67&ID%in%early.HFD))

null.lm.lme<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = LM.data.all)
age.lm.lme<-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = LM.data.all)
anova(null.lm.lme, age.lm.lme)#significant effect of age
diet.lm.lme<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = LM.data.all)
anova(age.lm.lme, diet.lm.lme)%>%tidy()%>%kable()#p=0.09
sex.lm.lme<-lmer(gram.values~ age + treatment*sex +(1|MouseID) + (1|dam), data = LM.data.all)
anova(sex.lm.lme, diet.lm.lme)#p<0.0001
Anova(sex.lm.lme)%>%tidy()%>%kable()#p=0.448
```

```{r Fig2B}
#Body Fat Data table
FM.data<-filter(MRI.raw.data, assay=="Total Fat Mass")%>%
  filter(age<70)%>%
  filter(!(age>=67&ID%in%early.HFD))%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
#Body Fat Plot
ggplot(FM.data,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 1,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Fat Mass", y = "Fat Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=18),legend.position = c(0.9,0.80), legend.background = element_rect(fill = "transparent"))
#Stats
FM.data.all<-MRI.raw.data%>%
  filter(assay=="Total Fat Mass")%>%
    filter(age<70)%>%
  filter(!(age>=67&ID%in%early.HFD))

null.fm.lme<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = FM.data.all)
age.fm.lme<-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = FM.data.all)
anova(null.fm.lme, age.fm.lme)#significant effect of age
sex.fm.lme<-lmer(gram.values~ age + treatment*sex +(1|MouseID) + (1|dam), data = FM.data.all)
diet.fm.lme<-lmer(gram.values ~ treatment +age +treatment:age+ (1|MouseID) + (1|dam), data = FM.data.all)
anova(sex.fm.lme, diet.fm.lme)#p=0.0000089
Anova(sex.bw.lme)%>%tidy()%>%kable()
```
#Post HFD analysis(PND 70-sacrifice)
```{r Fig3A}
#Body Weight data table
BW.data.HFD<-filter(MRI.raw.data, assay=="Body Weight")%>%
  filter(age>=70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

##Body Weight plot
ggplot(BW.data.HFD,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Body Weight", y = "Body weight (grams)",x="Age (days)")+
  theme(text = element_text(size=18),legend.position = c(0.9,0.19), legend.background = element_rect(fill = "transparent"))

#ALL
HFD.BW.data<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(assay=="Body Weight")


#mixed linear models
HFD.null.bw.lme<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.BW.data)
HFD.age.bw.lme <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.BW.data)
anova(HFD.null.bw.lme, HFD.age.bw.lme)#significant effect of age
HFD.diet.bw.lme<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.BW.data)
anova(HFD.age.bw.lme, HFD.diet.bw.lme)#sig with interaction
Anova(HFD.diet.bw.lme)
HFD.sex.bw.lme<-lmer(gram.values ~ age +sex +treatment + (1|MouseID) + (1|dam), data = HFD.BW.data)
anova(HFD.diet.bw.lme, HFD.sex.bw.lme)#significant, but no interaction between sex and treatment
Anova(HFD.sex.bw.lme)# no interaction of sex and treatment, so removed, 
```


```{r Fig3B}
#Fat Mass data table
FM.data.HFD<-filter(MRI.raw.data, assay=="Total Fat Mass")%>%
  filter(age>=70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

##Fat Mass plot
ggplot(FM.data.HFD,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values =color.scheme)+
  labs(title = "Offspring Fat Mass", y = "Fat Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=18),legend.position = c(0.9,0.19), legend.background = element_rect(fill = "transparent"))

#All
HFD.FM.data<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(assay=="Total Fat Mass")

HFD.null.fm.lme<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.FM.data)
HFD.age.fm.lme <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.FM.data)
anova(HFD.null.fm.lme, HFD.age.fm.lme)#significant effect of age
HFD.diet.fm.lme<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.FM.data)
anova(HFD.age.fm.lme, HFD.diet.fm.lme)#sig with interaction, p=0.00011
Anova(HFD.diet.fm.lme)
HFD.sex.fm.lme<-lmer(gram.values ~ age +sex*treatment + (1|MouseID) + (1|dam), data = HFD.FM.data)
anova(HFD.diet.fm.lme, HFD.sex.fm.lme)#significant, but no interaction between sex and treatment
Anova(HFD.sex.fm.lme)# no interaction of sex and treatment, so removed, 
```

```{r Fig3C}
#Lean Mass data table
LM.data.HFD<-filter(MRI.raw.data, assay=="Lean Mass")%>%
  filter(age>=70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

##Lean Mass plot
ggplot(LM.data.HFD,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Lean Mass", y = "Lean Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=18),legend.position = c(0.9,0.19), legend.background = element_rect(fill = "transparent"))
#stats
HFD.LM.data<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(assay=="Lean Mass")

HFD.null.lm.lme<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.LM.data)
HFD.age.lm.lme <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.LM.data)
anova(HFD.null.lm.lme, HFD.age.lm.lme)#significant effect of age
HFD.diet.lm.lme<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.LM.data)
anova(HFD.age.lm.lme, HFD.diet.lm.lme)# not sig with interaction, p=0.13
HFD.sex.lm.lme<-lmer(gram.values ~ age +sex*treatment + (1|MouseID) + (1|dam), data = HFD.LM.data)
anova(HFD.diet.lm.lme, HFD.sex.lm.lme)#significant, but no interaction between sex and treatment
Anova(HFD.sex.lm.lme)# no interaction of sex and treatment, so removed, 
```

#maternal counts
```{r maternal-info}
MRI.raw.data %>%
  group_by(dam,treatment) %>%
    distinct(dam) %>%
  group_by(treatment)%>%
  count(treatment) %>%
  kable(caption="Dams per feeding group")
```