---
title: "eTRF offspring body composition analysis"
author: "Molly C. Mulcahy"
date: "2/25/2021"
output:
  html_document: default
  pdf_document: default
---
#This document contains primary data and analysis for the dissertation work involving early time-restricted feeding during gestation effect on mouse offspring

#setup chunk
```{r setup, include=FALSE}
library(processx)
library(devtools)
library(dplyr)
library(readr)
library(car)
library(knitr)
library(tidyr)
library(lme4)
library(broom)
library(RCurl)
library(ggplot2)

#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively
knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})
#write standard error function for later use
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
# sets color scheme for plots
color.scheme <- c('gray69','gray36')
```

#Data Entry
```{r data-entry-MRI}
MRI.data.file<-"MRI.Off.data.csv"
MRI.raw.data<-read_csv(MRI.data.file,
                       col_types = cols(
  ID = col_factor(levels = NULL),
  age = col_double(),
  animal.id = col_factor(levels = NULL),
  MouseID = col_factor(levels = NULL),
  Sex = col_factor(levels = c("M","F")),
  Genotype = col_factor(levels = NULL),
  values = col_double(),
  assay = col_factor(levels = NULL),
  experiment.date = col_date(format = ""),
  treatment = col_factor(levels = c("AL","eTRF")),
  dam = col_factor(levels = NULL),
  cohort = col_factor(levels = NULL),
  gram.values = col_double()))

MRI.raw.data<-MRI.raw.data%>%
  mutate(sex = case_when(
    Sex =="M"~"male",
    Sex=="F"~"female"))
```
#Pre-HFD analysis (PND 21- PND 70)
#Body Composition
```{r Figure2A}
#table of unique animals for body composition numbers
MRI.raw.data %>%
  group_by(sex,treatment) %>%
  distinct(ID, .keep_all = T) %>%
  count %>%
  kable(caption="Animals in each group of this cohort")
#Body Weight data table
BW.data<-filter(MRI.raw.data, assay=="Body Weight")%>%
  filter(age<70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
##Body Weight plot
ggplot(BW.data,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Body Weight", y = "Body weight (grams)",x="Age (days)")+
  theme(text = element_text(size=16),legend.position = c(0.9,0.19))

#Females
#female body weight data set
BW.data.females<-MRI.raw.data%>%
  filter(age<70)%>%
  filter(sex=="female")%>%
  filter(assay=="Body Weight")
#mixed linear models
null.bw.lme.f<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = BW.data.females)
age.bw.lme.f <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = BW.data.females)
anova(null.bw.lme.f, age.bw.lme.f)#significant effect of age
diet.bw.lme.f<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = BW.data.females)
anova(age.bw.lme.f, diet.bw.lme.f)#significant diet age interaction, p=0.014
anova(diet.bw.lme.f)%>%tidy%>%kable()
#anova(diet.bw.lme.f)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
F.BW.effect<-fixef(diet.bw.lme.f)
#Males
#male body weight data set
BW.data.males<-MRI.raw.data%>%
  filter(age<70)%>%
  filter(sex=="male")%>%
  filter(assay=="Body Weight")

null.bw.lme.m<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = BW.data.males)
age.bw.lme.m <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = BW.data.males)
anova(null.bw.lme.m, age.bw.lme.m)#significant
diet.bw.lme.m<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = BW.data.males)
anova(age.bw.lme.m, diet.bw.lme.m)#NOT SIGNIFICANT, p=0.7
anova(diet.bw.lme.m)
fixef(diet.bw.lme.m)
```
There is a significant interaction of age and treatment in females where age-matched AL female have a ,p-value.

```{r Fig2C}
#Lean Mass data table
LM.data<-filter(MRI.raw.data, assay=="Lean Mass")%>%
  filter(age<70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
#Lean Mass Plot
ggplot(LM.data,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Lean Mass", y = "Lean Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=16),legend.position = c(0.9,0.80))

#stats

#Females
#female lean mass data set
LM.data.females<-MRI.raw.data%>%
  filter(assay=="Lean Mass")%>%
   filter(age<70)%>%
  filter(sex=="female")

null.lm.lme.f<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = LM.data.females)
age.lm.lme.f <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = LM.data.females)
anova(null.lm.lme.f, age.lm.lme.f)#significant effect of age
diet.lm.lme.f<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = LM.data.females)
anova(age.lm.lme.f, diet.lm.lme.f)#significant diet age interaction, p=0.0062
#anova(diet.lm.lme.f)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
fixef(diet.lm.lme.f)#
coefficients(diet.lm.lme.f)

#Males
#male body weight data set
LM.data.males<-MRI.raw.data%>%
  filter(assay=="Lean Mass")%>%
  filter(age<70)%>%
  filter(sex=="male")

null.lm.lme.m<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = LM.data.males)
age.lm.lme.m <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = LM.data.males)
anova(null.lm.lme.m, age.lm.lme.m)#significant
diet.lm.lme.m<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = LM.data.males)
anova(age.lm.lme.m, diet.lm.lme.m)#NOT SIGNIFICANT, p=0.75
#anova(diet.lm.lme.m)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
fixef(diet.lm.lme.m)#
coefficients(diet.lm.lme.m)
```

```{r Fig2B}
#Body Fat Data table
FM.data<-filter(MRI.raw.data, assay=="Total Fat Mass")%>%
  filter(age<70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
#Body Fat Plot
ggplot(FM.data,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme)+
  labs(title = "Offspring Fat Mass", y = "Fat Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=16),legend.position = c(0.9,0.80))
#Stats
#females
FM.data.females<-MRI.raw.data%>%
  filter(assay=="Total Fat Mass")%>%
  filter(age<70)%>%
  filter(sex=="female")

null.fm.lme.f<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = FM.data.females)
age.fm.lme.f <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = FM.data.females)
anova(null.fm.lme.f, age.fm.lme.f)#significant
diet.fm.lme.f<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = FM.data.females)
anova(age.fm.lme.f, diet.fm.lme.f)# p=0.43
#anova(diet.fm.lme.f)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
fixef(diet.fm.lme.f)#
coefficients(diet.fm.lme.f)
#males
FM.data.males<-MRI.raw.data%>%
  filter(assay=="Total Fat Mass")%>%
  filter(age<70)%>%
  filter(sex=="male")

null.fm.lme.m<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = FM.data.males)
age.fm.lme.m <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = FM.data.males)
anova(null.fm.lme.m, age.fm.lme.m)#significant
diet.fm.lme.m<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = FM.data.males)
anova(age.fm.lme.m, diet.fm.lme.m)# p=0.3
#anova(diet.fm.lme.m)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
fixef(diet.fm.lme.m)#
coefficients(diet.fm.lme.m)
```
#Food Intake Data
```{r data-entry-intake}
intake.data.raw<-("Intake.Off.data.csv")
intake.data<-read_csv(intake.data.raw,
                      col_types = cols(
                        cage = col_factor(levels = NULL),
                        "# in cage" = col_double(),
                        week = col_integer(),
                        "IDs" = col_factor(levels = NULL),
                        start = col_double(),
                        end=col_double(),
                        date = col_character(),
                        notes = col_factor(levels=NULL),
                        diet = col_factor(levels = NULL),
                        cohort = col_factor(levels = NULL),
                        food.consumed = col_double(),
                        kcals.consumed = col_double(),
                        food.cumulative = col_double(),
                        treatment = col_factor(levels = c("eTRF","AL")),
                        sex = col_factor(levels = c("female","male"))
                        ))
```

#Food Intake
```{r Fig2D}
#Table of Cages in Food Intake Analyses
intake.data %>%
  group_by(sex,treatment) %>%
  distinct(cage, .keep_all = T) %>%
  count %>%
  kable(caption="Cages in each group")

#make averaged dataset
average.intake<-intake.data%>%
   filter(!(week=="5"&cage=="2031"))%>%
   filter(!(week=="6"&cage=="2031"))%>%
   filter(!(week=="7"&cage=="2031"))%>%
  group_by(week, diet, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

#Filter out HFD from dataset
Pre.HFD.data<-average.intake%>%
  filter(diet=="NCD")

intake.data.NCD<-intake.data%>%
  filter(diet=="NCD")

#Plot of averaged.intake
ggplot(Pre.HFD.data, aes(week, kcals.consumed_mean,col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se=FALSE, span =0.7)+
  geom_errorbar(aes(ymin = kcals.consumed_mean -kcals.consumed_se, ymax = kcals.consumed_mean +kcals.consumed_se), width = 0.3)+
  facet_grid(.~sex)+
  labs(title="Weekly Food Intake", y = "Food Intake (kcals)", x="Week")+
  scale_color_manual(values = color.scheme)+
   theme(text = element_text(size=16),legend.position = c(0.1,0.80))


##Female
intake.data.NCD.F<-intake.data.NCD%>%
  filter(sex=="female")

null.week.food.lme.F<-lmer(kcals.consumed ~ 1 + (1|cage), intake.data.NCD.F)
age.week.food.lme.F<-lmer(kcals.consumed ~ week + (1|cage), intake.data.NCD.F)
anova(null.week.food.lme.F, age.week.food.lme.F)# no effect of age,p= 0.068
group.week.food.lme.F<-lmer(kcals.consumed~ week*treatment +(1|cage), intake.data.NCD.F)
anova(age.week.food.lme.F, group.week.food.lme.F)# effect of group in female(p=0.049)
fixef(group.week.food.lme.F)
##Male
intake.data.NCD.M<-intake.data.NCD%>%
  filter(sex=="male")

null.week.food.lme.M<-lmer(kcals.consumed ~ 1 + (1|cage), intake.data.NCD.M)
age.week.food.lme.M<-lmer(kcals.consumed ~ week + (1|cage), intake.data.NCD.M)
anova(null.week.food.lme.M, age.week.food.lme.M)# effect of age,p= 0.00079
group.week.food.lme.M<-lmer(kcals.consumed~ week*treatment +(1|cage), intake.data.NCD.M)
anova(age.week.food.lme.M, group.week.food.lme.M)# no effect of group in male(p=0.60)
fixef(group.week.food.lme.M)
```


```{r fig2E}
#cumulative intake plot
ggplot(Pre.HFD.data, aes(week, food.cumulative_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se=FALSE)+
  geom_errorbar(aes(ymin = food.cumulative_mean -food.cumulative_se, ymax = food.cumulative_mean +food.cumulative_se), width = 0.3)+
  facet_grid(.~sex)+
  labs(x="Week",y="Cumulative Food Intake (kcals)")+
  ggtitle("Cumulative Food Intake")+
  scale_color_manual(values = color.scheme)+
  theme(text = element_text(size=16),legend.position = c(0.1,0.80))



#statistics for cumulative food intake

##Female
null.cumu.food.lme.F<-lmer(food.cumulative ~ 1 + (1|cage), intake.data.NCD.F)
age.cumu.food.lme.F<-lmer(food.cumulative~ week + (1|cage), intake.data.NCD.F)
anova(null.cumu.food.lme.F, age.cumu.food.lme.F)# effect of age,p< 0.0001
group.cumu.food.lme.F<-lmer(food.cumulative~ week*treatment +(1|cage), intake.data.NCD.F)
anova(age.cumu.food.lme.F, group.cumu.food.lme.F)#effect of group in female (p=0.00022)
fixef(group.cumu.food.lme.F)
##Male
null.cumu.food.lme.M<-lmer(food.cumulative ~ 1 + (1|cage), intake.data.NCD.M)
age.cumu.food.lme.M<-lmer(food.cumulative~ week + (1|cage), intake.data.NCD.M)
anova(null.cumu.food.lme.M, age.cumu.food.lme.M)# effect of age,p< 0.0001
group.cumu.food.lme.M<-lmer(food.cumulative~ week*treatment +(1|cage), intake.data.NCD.M)
anova(age.cumu.food.lme.M, group.cumu.food.lme.M)# no effect of group in male(p=0.34)
fixef(group.cumu.food.lme.M)

```

```{r Fig2F}
NCD.ITT.raw.data<-read_csv("NCD.ITT.Offspring.csv",
                           col_types = cols(
                             ID = col_factor(levels = NULL),
                             cage = col_factor(levels = NULL),
                             time = col_integer(),
                             glucose = col_double(),
                             treatment = col_factor(levels = c("eTRF","AL")),
                             sex = col_factor(levels = c("female","male")),
                             cohort = col_factor(levels = NULL),
                             dam = col_factor(levels = NULL),
                             rel.glucose = col_double()
                           ))

summary.ITT<-NCD.ITT.raw.data%>%
  group_by(sex, time, treatment)%>%
  summarise_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

#plot, time course
ggplot(summary.ITT, aes(time, glucose_mean))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se = FALSE)+
  geom_errorbar(aes(ymin = glucose_mean -glucose_se, ymax = glucose_mean + glucose_se))+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme )+
  labs(title = "Insulin Tolerance Test", y = "Blood Glucose (mg/dL)", x = "Time (minutes)")+
   theme(text = element_text(size=16),legend.position = c(0.9,0.80))

#Stats
##Males
NCD.ITT.M<-NCD.ITT.raw.data%>%
  filter(sex == "male")
null.NCDITT.lme.m<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = NCD.ITT.M)
age.NCDITT.lme.m <-lmer(glucose ~ time + (1|ID) + (1|dam), data = NCD.ITT.M)
anova(null.NCDITT.lme.m, age.NCDITT.lme.m)#significant, p=0.0062
diet.NCDITT.lme.m<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = NCD.ITT.M)
anova(age.NCDITT.lme.m, diet.NCDITT.lme.m)# p=0.062
fixef(diet.NCDITT.lme.m)
##Females
NCD.ITT.F<-NCD.ITT.raw.data%>%
  filter(sex == "female")
null.NCDITT.lme.f<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = NCD.ITT.F)
age.NCDITT.lme.f <-lmer(glucose ~ time + (1|ID) + (1|dam), data = NCD.ITT.F)
anova(null.NCDITT.lme.f, age.NCDITT.lme.f)#significant, p=0.0021
diet.NCDITT.lme.f<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = NCD.ITT.F)
anova(age.NCDITT.lme.f, diet.NCDITT.lme.f)# p=0.88
fixef(diet.NCDITT.lme.m)
```

```{r Fig2G}
   #FBG
  FBG.data<- NCD.ITT.raw.data%>%
     filter(time=="0")%>%
    group_by(sex, treatment)%>%
    summarise_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))
  #FBG plot
  ggplot(FBG.data, aes(treatment, glucose_mean, fill = treatment))+
    geom_col(aes(fill = treatment))+
    facet_grid(.~sex)+
    geom_errorbar(aes(ymin = glucose_mean -glucose_se, ymax = glucose_mean + glucose_se), width = 0.3)+
    scale_fill_manual(values = color.scheme)+
    labs(title = "Fasting Blood Glucose", y = "Blood Glucose (mg/dL)")+
   theme(text = element_text(size=16))
  
  #stats
  ##female
  FBG.stats.F<-NCD.ITT.raw.data%>%
     filter(time=="0")%>%
    filter(sex=="female")
leveneTest(glucose~treatment, FBG.stats.F)#p=0.45, equal variance
shapiro.test(FBG.stats.F$glucose[FBG.stats.F$treatment=="AL"])#0.5, normal
shapiro.test(FBG.stats.F$glucose[FBG.stats.F$treatment=="eTRF"])#0.6, normal

t.test(glucose~ treatment, data = FBG.stats.F)#p=0.3
  ##male
  FBG.stats.M<-NCD.ITT.raw.data%>%
     filter(time=="0")%>%
    filter(sex=="male")
leveneTest(glucose~treatment, FBG.stats.M)#p=0.92, equal variance
shapiro.test(FBG.stats.M$glucose[FBG.stats.M$treatment=="AL"])#0.6, normal
shapiro.test(FBG.stats.M$glucose[FBG.stats.M$treatment=="eTRF"])#0.9, normal

t.test(glucose~ treatment, data = FBG.stats.M)#p=0.2
  
```

```{r Fig2H}
NCD.GTT.data.file<-"NCD.GTT.raw.data.csv"
NCD.GTT.raw.data<-read_csv(NCD.GTT.data.file,
    col_types = cols(
  ID = col_factor(levels = NULL),
  time = col_integer(),
  glucose = col_double(),
  cage = col_factor(levels = NULL),
  dam = col_factor(levels = NULL),
  sex = col_factor(levels = NULL),
  treatment = col_factor(levels = c("eTRF","AL")),
  cohort = col_factor(levels = NULL)
  ))
#Summary data set
summary.GTT<-NCD.GTT.raw.data%>%
  group_by(sex, time, treatment)%>%
  summarise_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

NCD.GTT.raw.data %>%
  group_by(sex,treatment) %>%
  distinct(ID, .keep_all = T) %>%
  count %>%
  kable(caption="IDS in each group")

#plot, time course
ggplot(summary.GTT, aes(time, mean))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se = FALSE)+
  geom_errorbar(aes(ymin = mean -se, ymax = mean + se))+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme )+
  labs(title = "Glucose Tolerance Test", y = "Blood Glucose (mg/dL)", x = "Time (minutes)")+
   theme(text = element_text(size=16),legend.position = c(0.4,0.80))

#Stats
##Males
NCD.GTT.M<-NCD.GTT.raw.data%>%
  filter(sex == "male")
null.NCDGTT.lme.m<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = NCD.GTT.M)
age.NCDGTT.lme.m <-lmer(glucose ~ time + (1|ID) + (1|dam), data = NCD.GTT.M)
anova(null.NCDGTT.lme.m, age.NCDGTT.lme.m)#significant, p=0.0062
diet.NCDGTT.lme.m<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = NCD.GTT.M)
anova(age.NCDGTT.lme.m, diet.NCDGTT.lme.m)# p=0.17
fixef(diet.NCDGTT.lme.m)
##Females
NCD.ITT.F<-NCD.GTT.raw.data%>%
  filter(sex == "female")
null.NCDITT.lme.f<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = NCD.ITT.F)
age.NCDITT.lme.f <-lmer(glucose ~ time + (1|ID) + (1|dam), data = NCD.ITT.F)
anova(null.NCDITT.lme.f, age.NCDITT.lme.f)#significant, p=0.0021
diet.NCDITT.lme.f<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = NCD.ITT.F)
anova(age.NCDITT.lme.f, diet.NCDITT.lme.f)# p=0.79
fixef(diet.NCDITT.lme.m)
```
#Post HFD analysis(PND 70-sacrifice)

```{r Fig3A}
#Body Weight data table
BW.data.HFD<-filter(MRI.raw.data, assay=="Body Weight")%>%
  filter(age>=70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

##Body Weight plot
ggplot(BW.data.HFD,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = c("gray36", "gray69"))+
  labs(title = "Offspring Body Weight", y = "Body weight (grams)",x="Age (days)")+
  theme(text = element_text(size=16),legend.position = c(0.9,0.19))

#Females
#female body weight data set
HFD.BW.data.females<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(sex=="female")%>%
  filter(assay=="Body Weight")
#mixed linear models
HFD.null.bw.lme.f<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.BW.data.females)
HFD.age.bw.lme.f <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.BW.data.females)
anova(HFD.null.bw.lme.f, HFD.age.bw.lme.f)#significant effect of age
HFD.diet.bw.lme.f<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.BW.data.females)
anova(HFD.age.bw.lme.f, HFD.diet.bw.lme.f)#no diet age interaction, p=0.073
#anova(HFD.diet.bw.lme.f)%>%tidy%>%kable()
#anova(diet.bw.lme.f)%>%tidy%>%filter(term =="age:treatment")%>%pull(p.value)
F.BW.effect<-fixef(HFD.diet.bw.lme.f)
#Males
#male body weight data set
BW.data.males<-MRI.raw.data%>%
  filter(age<70)%>%
  filter(Sex=="M")%>%
  filter(assay=="Body Weight")

null.bw.lme.m<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = BW.data.males)
age.bw.lme.m <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = BW.data.males)
anova(null.bw.lme.m, age.bw.lme.m)#significant
diet.bw.lme.m<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = BW.data.males)
anova(age.bw.lme.m, diet.bw.lme.m)#NOT SIGNIFICANT, p=0.7
anova(diet.bw.lme.m)
fixef(diet.bw.lme.m)

```


```{r Fig3B}
#Fat Mass data table
FM.data.HFD<-filter(MRI.raw.data, assay=="Total Fat Mass")%>%
  filter(age>=70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

##Fat Mass plot
ggplot(FM.data.HFD,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = c("gray36", "gray69"))+
  labs(title = "Offspring Fat Mass", y = "Fat Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=16),legend.position = c(0.9,0.19))

#Females
#female fat mass data set
HFD.FM.data.females<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(sex=="female")%>%
  filter(assay=="Total Fat Mass")
#mixed linear models
HFD.null.fm.lme.f<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.FM.data.females)
HFD.age.fm.lme.f <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.FM.data.females)
anova(HFD.null.fm.lme.f, HFD.age.fm.lme.f)#significant effect of age
HFD.diet.fm.lme.f<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.FM.data.females)
anova(HFD.age.fm.lme.f, HFD.diet.fm.lme.f)#p=0.30
fixef(HFD.diet.fm.lme.f)
##males
#male fat mass data set
HFD.FM.data.males<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(sex=="male")%>%
  filter(assay=="Total Fat Mass")
#mixed linear models
HFD.null.fm.lme.m<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.FM.data.males)
HFD.age.fm.lme.m <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.FM.data.males)
anova(HFD.null.fm.lme.m, HFD.age.fm.lme.m)#significant effect of age
HFD.diet.fm.lme.m<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.FM.data.males)
anova(HFD.age.fm.lme.m, HFD.diet.fm.lme.m)#p=0.075
fixef(HFD.diet.fm.lme.m)
```

```{r Fig3C}
#Lean Mass data table
LM.data.HFD<-filter(MRI.raw.data, assay=="Lean Mass")%>%
  filter(age>=70)%>%
  group_by(age, treatment, sex)%>%
  summarize_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

##Lean Mass plot
ggplot(LM.data.HFD,aes(age, gram.values_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(span = 0.7,aes(col = treatment), se = FALSE)+
  facet_grid(.~sex)+
  scale_color_manual(values = c("gray36", "gray69"))+
  labs(title = "Offspring Lean Mass", y = "Lean Mass (grams)",x="Age (days)")+
  theme(text = element_text(size=16),legend.position = c(0.9,0.19))
#stats
##females
#female fat mass data set
HFD.LM.data.females<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(sex=="female")%>%
  filter(assay=="Lean Mass")
#mixed linear models
HFD.null.lm.lme.f<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.LM.data.females)
HFD.age.lm.lme.f <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.LM.data.females)
anova(HFD.null.lm.lme.f, HFD.age.lm.lme.f)#significant effect of age
HFD.diet.lm.lme.f<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.LM.data.females)
anova(HFD.age.lm.lme.f, HFD.diet.lm.lme.f)#p=0.85
fixef(HFD.diet.lm.lme.f)
##males
#male fat mass data set
HFD.LM.data.males<-MRI.raw.data%>%
  filter(age>=70)%>%
  filter(sex=="male")%>%
  filter(assay=="Lean Mass")
#mixed linear models
HFD.null.lm.lme.m<-lmer(gram.values ~ 1 + (1|MouseID) + (1|dam), data = HFD.LM.data.males)
HFD.age.lm.lme.m <-lmer(gram.values ~ age + (1|MouseID) + (1|dam), data = HFD.LM.data.males)
anova(HFD.null.lm.lme.m, HFD.age.lm.lme.m)#significant effect of age
HFD.diet.lm.lme.m<-lmer(gram.values ~ treatment+age +age:treatment + (1|MouseID) + (1|dam), data = HFD.LM.data.males)
anova(HFD.age.lm.lme.m, HFD.diet.lm.lme.m)#p=0.28
fixef(HFD.diet.lm.lme.m)
```

```{r Fig3D}
#Filter out NCD from dataset
HFD.data<-average.intake%>%
  filter(diet=="HFD")

intake.data.HFD<-intake.data%>%
  filter(diet=="HFD")

#Plot of averaged.intake
ggplot(HFD.data, aes(week, kcals.consumed_mean,col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se=FALSE, span =0.7)+
  geom_errorbar(aes(ymin = kcals.consumed_mean -kcals.consumed_se, ymax = kcals.consumed_mean +kcals.consumed_se), width = 0.3)+
  facet_grid(.~sex)+
  labs(title="Weekly Food Intake", y = "Food Intake (kcals)", x="Week")+
  scale_color_manual(values = color.scheme)+
   theme(text = element_text(size=16),legend.position = c(0.9,0.80))

#stats
intake.data.HFD.F<-intake.data.HFD%>%
  filter(sex=="female")

null.week.HFD.food.lme.F<-lmer(kcals.consumed ~ 1 + (1|cage), intake.data.HFD.F)
age.week.HFD.food.lme.F<-lmer(kcals.consumed ~ week + (1|cage), intake.data.HFD.F)
anova(null.week.HFD.food.lme.F, age.week.HFD.food.lme.F)# no effect of age,p= 0.068
group.week.HFD.food.lme.F<-lmer(kcals.consumed~ week*treatment +(1|cage), intake.data.HFD.F)
anova(age.week.HFD.food.lme.F, group.week.HFD.food.lme.F)# no effect of group in female(p=0.76)
fixef(group.week.HFD.food.lme.F)
##Male
intake.data.HFD.M<-intake.data.HFD%>%
  filter(sex=="male")

null.week.HFD.food.lme.M<-lmer(kcals.consumed ~ 1 + (1|cage), intake.data.HFD.M)
age.week.HFD.food.lme.M<-lmer(kcals.consumed ~ week + (1|cage), intake.data.HFD.M)
anova(null.week.HFD.food.lme.M, age.week.HFD.food.lme.M)# no effect of age,p= 0.068
group.week.HFD.food.lme.M<-lmer(kcals.consumed~ week*treatment +(1|cage), intake.data.HFD.M)
anova(age.week.HFD.food.lme.M, group.week.HFD.food.lme.M)# no effect of group in female(p=0.57)
fixef(group.week.HFD.food.lme.M)
```


```{r Fig3E}

HFD.data.filtered<-HFD.data%>%
  filter(!(week == 15))
#Plot of  cumulative intake
ggplot(HFD.data.filtered, aes(week, food.cumulative_mean,col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se=FALSE, span =0.7)+
  geom_errorbar(aes(ymin = food.cumulative_mean -food.cumulative_se, ymax = food.cumulative_mean +food.cumulative_se), width = 0.3)+
  facet_grid(.~sex)+
  labs(title="Cumulative Food Intake", y = "Cumulative Food Intake (kcals)", x="Week")+
  scale_color_manual(values = color.scheme)+
   theme(text = element_text(size=16),legend.position = c(0.1,0.80))

#statistics
##Female
null.cumu.HFD.food.lme.F<-lmer(food.cumulative ~ 1 + (1|cage), intake.data.HFD.F)
age.cumu.HFD.food.lme.F<-lmer(food.cumulative~ week + (1|cage), intake.data.HFD.F)
anova(null.cumu.HFD.food.lme.F, age.cumu.HFD.food.lme.F)# effect of age,p< 0.0001
group.cumu.HFD.food.lme.F<-lmer(food.cumulative~ week*treatment +(1|cage), intake.data.HFD.F)
anova(age.cumu.HFD.food.lme.F, group.cumu.HFD.food.lme.F)# noeffect of group in female (p=0.47)
fixef(group.cumu.HFD.food.lme.F)
##Male
null.cumu.HFD.food.lme.M<-lmer(food.cumulative ~ 1 + (1|cage), intake.data.HFD.M)
age.cumu.HFD.food.lme.M<-lmer(food.cumulative~ week + (1|cage), intake.data.HFD.M)
anova(null.cumu.HFD.food.lme.M, age.cumu.HFD.food.lme.M)# effect of age,p< 0.0001
group.cumu.HFD.food.lme.M<-lmer(food.cumulative~ week*treatment +(1|cage), intake.data.HFD.M)
anova(age.cumu.HFD.food.lme.M, group.cumu.HFD.food.lme.M)# no effect of group in male(p=0.55)
fixef(group.cumu.HFD.food.lme.M)

```


```{r Fig3F}
#High fat diet Insulin tolerance test
HFD.ITT.file <-"HFD.ITT.Offspring.csv"
HFD.ITT.raw.data<-read_csv(HFD.ITT.file,
                           col_types = cols(
                             cage = col_factor(levels = NULL),
                               ID = col_factor(levels = NULL),
                               time = col_integer(),
                               glucose = col_double(),
                               crash = col_factor(levels = NULL),
                               dam = col_factor(levels = NULL),
                               cohort = col_factor(levels = NULL),
                               treatment = col_factor(levels = c("eTRF","AL")),
                               sex = col_factor(levels = c("female","male")),
                               rel.glucose = col_double()
                           ))

#Summary data set
summary.HFD.ITT<-HFD.ITT.raw.data%>%
  group_by(sex, time, treatment)%>%
  summarise_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

HFD.ITT.raw.data %>%
  group_by(sex,treatment) %>%
  distinct(ID, .keep_all = T) %>%
  count %>%
  kable(caption="IDS in each group")

#plot, time course
ggplot(summary.HFD.ITT, aes(time, glucose_mean))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se = FALSE)+
  geom_errorbar(aes(ymin = glucose_mean - glucose_se, ymax = glucose_mean + glucose_se))+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme )+
  labs(title = "Insulin Tolerance Test", y = "Blood Glucose (mg/dL)", x = "Time (minutes)")+
   theme(text = element_text(size=16),legend.position = c(0.4,0.80))

#Stats
##Males
HFD.ITT.M<-HFD.ITT.raw.data%>%
  filter(sex == "male")
null.HFD.ITT.lme.m<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = HFD.ITT.M)
age.HFD.ITT.lme.m <-lmer(glucose ~ time + (1|ID) + (1|dam), data = HFD.ITT.M)
anova(null.HFD.ITT.lme.m, age.HFD.ITT.lme.m)#significant, p=0.0062
diet.HFD.ITT.lme.m<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = HFD.ITT.M)
anova(age.HFD.ITT.lme.m, diet.HFD.ITT.lme.m)# p=0.17
fixef(diet.HFD.ITT.lme.m)
##females
##Males
HFD.ITT.F<-HFD.ITT.raw.data%>%
  filter(sex == "female")
null.HFD.ITT.lme.f<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = HFD.ITT.F)
age.HFD.ITT.lme.f <-lmer(glucose ~ time + (1|ID) + (1|dam), data = HFD.ITT.F)
anova(null.HFD.ITT.lme.f, age.HFD.ITT.lme.f)#significant, p=0.0062
diet.HFD.ITT.lme.f<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = HFD.ITT.F)
anova(age.HFD.ITT.lme.f, diet.HFD.ITT.lme.f)# p=0.85
fixef(diet.HFD.ITT.lme.f)
```

```{r Fig3G}
#generate HFD ITT data set
HFD.FBG.data<-HFD.ITT.raw.data%>%
  filter(time=="0")
#HFD ITT summary data set
summary.HFD.FBG<-HFD.FBG.data%>%
  group_by(sex, treatment)%>%
  summarise_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

#plot
ggplot(summary.HFD.FBG,  aes(treatment, glucose_mean, fill = treatment))+
  geom_col(aes(fill = treatment))+
  facet_grid(.~sex)+
  scale_fill_manual(values = color.scheme)+
  geom_errorbar(aes(ymin = glucose_mean - glucose_se, ymax = glucose_mean + glucose_se), width = 0.3)+
  labs(title = "Fasting Blood Glucose", y = "Blood Glucose (mg/dL)")+
   theme(text = element_text(size=16))

 #stats
  ##female
  HFD.FBG.stats.F<-HFD.ITT.raw.data%>%
     filter(time=="0")%>%
    filter(sex=="female")
leveneTest(glucose~treatment, HFD.FBG.stats.F)#p=0.76, equal variance
shapiro.test(HFD.FBG.stats.F$glucose[HFD.FBG.stats.F$treatment=="AL"])#0.6, normal
shapiro.test(HFD.FBG.stats.F$glucose[HFD.FBG.stats.F$treatment=="eTRF"])#0.99, normal

t.test(glucose~ treatment, data = HFD.FBG.stats.F)#p=0.2
  ##male
  HFD.FBG.stats.M<-HFD.ITT.raw.data%>%
     filter(time=="0")%>%
    filter(sex=="male")
leveneTest(glucose~treatment, HFD.FBG.stats.M)#p=0.34, equal variance
shapiro.test(HFD.FBG.stats.M$glucose[HFD.FBG.stats.M$treatment=="AL"])#0.9, normal
shapiro.test(HFD.FBG.stats.M$glucose[HFD.FBG.stats.M$treatment=="eTRF"])#0.3, normal

t.test(glucose~ treatment, data = HFD.FBG.stats.M)#p=0.4
```


```{r Fig3H}
#Glucose tolerance test
HFD.GTT.file<-"HFD.GTT.Offspring.csv"
HFD.GTT.raw.data<-read_csv(HFD.GTT.file,
                           col_types = cols(
                             cage = col_factor(levels = NULL),
                               ID = col_factor(levels = NULL),
                               time = col_integer(),
                               glucose = col_double(),
                               dam = col_factor(levels = NULL),
                               cohort = col_factor(levels = NULL),
                               treatment = col_factor(levels = c("eTRF","AL")),
                               sex = col_factor(levels = c("female","male")),
                               rel.glucose = col_double()
                           ))

##summary dataset
summary.HFD.GTT<-HFD.GTT.raw.data%>%
  group_by(sex, treatment, time)%>%
  summarise_if(is.numeric, .funs = funs(mean=mean(.,na.rm = TRUE),se=se, n=length))

#Plot
ggplot(summary.HFD.GTT, aes(time, glucose_mean, col = treatment))+
  geom_point(aes(col = treatment))+
  geom_smooth(aes(col = treatment), se = FALSE)+
  geom_errorbar(aes(ymin = glucose_mean - glucose_se, ymax = glucose_mean + glucose_se))+
  facet_grid(.~sex)+
  scale_color_manual(values = color.scheme )+
  labs(title = "Glucose Tolerance Test", y = "Blood Glucose (mg/dL)", x = "Time (minutes)")+
   theme(text = element_text(size=16),legend.position = c(0.4,0.80))

#Stats
##Males
HFD.GTT.M<-HFD.GTT.raw.data%>%
  filter(sex == "male")
null.HFD.GTT.lme.m<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = HFD.GTT.M)
age.HFD.GTT.lme.m <-lmer(glucose ~ time + (1|ID) + (1|dam), data = HFD.GTT.M)
anova(null.HFD.GTT.lme.m, age.HFD.GTT.lme.m)#significant, p=0.0062
diet.HFD.GTT.lme.m<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = HFD.GTT.M)
anova(age.HFD.GTT.lme.m, diet.HFD.GTT.lme.m)# p=0.14
fixef(diet.HFD.GTT.lme.m)
##females
HFD.GTT.F<-HFD.GTT.raw.data%>%
  filter(sex == "female")
null.HFD.GTT.lme.f<-lmer(glucose ~ 1 + (1|ID) + (1|dam), data = HFD.GTT.F)
age.HFD.GTT.lme.f <-lmer(glucose ~ time + (1|ID) + (1|dam), data = HFD.GTT.F)
anova(null.HFD.GTT.lme.f, age.HFD.GTT.lme.f)#significant, p=0.0062
diet.HFD.GTT.lme.f<-lmer(glucose ~ time + treatment + time:treatment + (1|ID) + (1|dam), data = HFD.GTT.F)
anova(age.HFD.GTT.lme.f, diet.HFD.GTT.lme.f)# p=0.61
fixef(diet.HFD.GTT.lme.f)
```
#maternal counts
```{r maternal-info}
MRI.raw.data %>%
  group_by(dam,treatment) %>%
    distinct(dam) %>%
  group_by(treatment)%>%
  count(treatment) %>%
  kable(caption="Dams per feeding group")
```