---
title: "dam.foodintake"
author: "Molly C. Mulcahy"
date: '2022-05-25'
output:
  html_document:
    toc: yes
    keep_md: yes
---

```{r setup}
#load commonly needed packages
library(processx)
library(devtools)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(knitr)
library(car)
library(lme4)
library(broom)
library(forcats)

#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively
knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))

#set plot color scheme
color.scheme = c("#333333","#CCCCCC")
#defense.color.scheme<- c("#00274C","#A5A508")
```

#data entry
```{r data-entry}
dam.foodintake.file<-"dam_foodintake.csv"

intake.data<-read_csv(dam.foodintake.file, 
  col_types = cols(
  MouseID = col_factor(levels = NULL),
  "food start" = col_double(),
  "food end" = col_double(),
  "number in cage" = col_double(),
  kcals.consumed = col_double(),
  Genotype = col_factor(levels = c("+/+","-/-")),
  gest.age = col_double(),
  postnatal = col_factor(levels = c("n","y")),
  food.cumulative = col_double()
  ))
          
intake.data.cleaned<-intake.data%>%
mutate( week = cut(gest.age, breaks = c(-7.5,0,7.5, 14.5, 21.5, 28.5, 35.5,42.5), labels = c("-1","0","1","2","3","4","5")))%>%
  mutate(week=as.numeric(week))
```


```{r prenatal-food-bar}
 #generate final prenatal food value for each animal
prenatal.food.cumulative<-intake.data.cleaned%>%
  filter(postnatal =="n")%>%
  arrange(week)%>%
  group_by(MouseID)%>%
  mutate(final.pre = last(food.cumulative))
#select only needed variables, one observation per animal
prenatal.final<-prenatal.food.cumulative%>%
  group_by(MouseID, Genotype)%>%
  distinct(MouseID, .keep_all = T)%>%
  select(MouseID, final.pre, Genotype)

#bar graph, prenatal food intake
prenatal.final%>%
  group_by(Genotype)%>%
  summarize(avg.kcal = mean(final.pre),
               error.kcal = se(final.pre))%>%
  ggplot(aes(Genotype, avg.kcal, fill = Genotype))+
  geom_col()+
  geom_errorbar(aes(ymin = avg.kcal -error.kcal, ymax = avg.kcal + error.kcal),width=0.3)+
  scale_fill_manual(values = color.scheme)+
  labs(title = "Prenatal cumulative food intake", y="Prenatal cumulative food intake (kcal/dam)")+
  theme_bw()+
  theme(legend.position = "none",text = element_text(size=14))

#prenatal food intake stats
shapiro.test(prenatal.final$final.pre[prenatal.final$Genotype=="-/-"])#0.5
shapiro.test(prenatal.final$final.pre[prenatal.final$Genotype=="+/+"])#0.08
leveneTest(final.pre ~ Genotype, data = prenatal.final)#0.98
t.test(final.pre ~ Genotype, data = prenatal.final)$p.value#p=0.5
```


```{r postnatal-food-bar}
 #generate final postnatal food value for each animal
postnatal.food.cumulative<-intake.data.cleaned%>%
  filter(postnatal =="y")%>%
  arrange(week)%>%
  group_by(MouseID)%>%
  mutate(final.post = last(food.cumulative))
#select only needed variables, one observation per animal
postnatal.final<-postnatal.food.cumulative%>%
  group_by(MouseID, Genotype)%>%
  distinct(MouseID, .keep_all = T)%>%
  select(MouseID, final.post, Genotype)
merged.final<-merge(postnatal.final, prenatal.final, by =c("Genotype", "MouseID"))
#merge data from prenatal period, subtract kcals from prenatal period
sub.post<-merged.final%>%
  group_by(MouseID)%>%
  mutate(postnatal.final = final.post-final.pre)

#bar graph, postnatal food intake
sub.post%>%
   group_by(Genotype)%>%
  summarize(avg.kcal = mean(postnatal.final),
               error.kcal = se(postnatal.final))%>%
  ggplot(aes(Genotype, avg.kcal, fill = Genotype))+
  geom_col()+
  geom_errorbar(aes(ymin = avg.kcal -error.kcal, ymax = avg.kcal + error.kcal),width=0.3)+
  scale_fill_manual(values = color.scheme)+
  labs(title = "Postnatal cumulative food intake", y="Postnatal cumulative food intake (kcal/dam)")+
  theme_bw()+
    theme(legend.position = "none",text = element_text(size=14))

#postnatal food intake stats
shapiro.test(sub.post$postnatal.final[sub.post$Genotype=="+/+"])#0.1
shapiro.test(sub.post$postnatal.final[sub.post$Genotype=="-/-"])#0.3
leveneTest(postnatal.final ~ Genotype, sub.post)#0.83
t.test(postnatal.final ~ Genotype, sub.post)$p.value#0.9
  
```
 
```{r weekly-plots}
 #Daily Food 
ggplot(intake.data.cleaned, aes(as.integer(week), kcals.consumed, col = Genotype))+
  geom_point(aes(col = Genotype))+
  geom_smooth(aes(col = Genotype), se = FALSE, span =0.5)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7),labels = c("pre","0","1","2","3","4","5"))+
  scale_color_manual(values = color.scheme)+
  labs(title = "Maternal Food Intake", y= "food intake (kcal/mouse/day)",x="Gestational Age")+
  theme_bw()+
  theme(legend.position = c(0.1,0.85), legend.background = element_blank(),text = element_text(size=14))

#Cumulative Food Intake
 intake.data.cleaned%>%
  #filter(postnatal=="n")%>%
   ggplot(aes(gest.age, food.cumulative, col = Genotype))+
   geom_point(aes(col = Genotype))+
    scale_color_manual(values = color.scheme)+
   geom_smooth(aes(col = Genotype),se=FALSE)+
  labs(title = "Maternal Cumulative Food Intake", y= "kcals/mouse",x="Gestational Day")+
  theme(legend.position = c(0.1,0.85),legend.background = element_blank())
```


```{ stats}
#pre birth
##daily kcals
pre.birth.intake<-intake.data.cleaned%>%
  filter(postnatal == "n")#filter for only pre-birth observations
daily.null<-lmer(kcals.consumed ~ 1 + (1|MouseID), data = pre.birth.intake)
daily.day<-lmer(kcals.consumed ~ as.factor(gest.age) + (1|MouseID), pre.birth.intake)
anova(daily.null, daily.day) #significant effect of gestational age
daily.geno<-lmer(kcals.consumed ~ gest.age + Genotype +(1|MouseID), pre.birth.intake)
anova(daily.day, daily.geno)#no significant effect for genotype
daily.geno.int<-lmer(kcals.consumed ~ gest.age*Genotype + (1|MouseID), pre.birth.intake)
anova(daily.geno, daily.geno.int)
Anova(daily.geno.int)# No interaction effect 
Anova(daily.geno)
fixef(daily.geno)

#cumulative kcals
cumu.null<-lmer(food.cumulative ~1 + (1|MouseID), data = pre.birth.intake)
cumu.day<-lmer(food.cumulative ~gest.age + (1|MouseID), data = pre.birth.intake)
anova(cumu.null, cumu.day)#sig effect of day
cumu.geno <- lmer(food.cumulative ~gest.age + Genotype + (1|MouseID), data = pre.birth.intake)
anova(cumu.day, cumu.geno)#sig efect of geno
cumu.geno.int <- lmer(food.cumulative ~gest.age*Genotype + (1|MouseID), data = pre.birth.intake)
anova(cumu.geno, cumu.geno.int)#no interaction
Anova(cumu.geno.int)# no interaction present, use simple model 
Anova(cumu.geno)
fixef(cumu.geno)


#post models
 post.birth.intake<-intake.data.cleaned%>%
  filter(postnatal=="y")
#daily kcals
post.null<-lmer(kcals.consumed ~ 1 + (1|MouseID), data = post.birth.intake)
daily.post<-lmer(kcals.consumed ~ gest.age + (1|MouseID), data = post.birth.intake)
anova(post.null, daily.post)#sig effect of day
geno.post<-lmer(kcals.consumed ~ gest.age + Genotype + (1|MouseID), data = post.birth.intake)
anova(daily.post, geno.post)#no significant difference between models
geno.post.int<-lmer(kcals.consumed ~ gest.age*Genotype + (1|MouseID), data = post.birth.intake)
anova(geno.post, geno.post.int)
Anova(geno.post.int)#no effect of interaction
Anova(geno.post)
fixef(geno.post)

#cumulative kcals
post.cumu.null<-lmer(food.cumulative ~ 1 + (1|MouseID), data = post.birth.intake)
daily.cumu.post<-lmer(food.cumulative ~ gest.age + (1|MouseID), data = post.birth.intake)
anova(post.cumu.null, daily.cumu.post)#significant effect of gestational age
geno.cumu.post<-lmer(food.cumulative ~ gest.age + Genotype + (1|MouseID), data = post.birth.intake)
anova(daily.cumu.post, geno.cumu.post)#significant difference between models
geno.int.cumu.post<-lmer(food.cumulative ~ gest.age*Genotype + (1|MouseID), data = post.birth.intake)
anova(geno.cumu.post, geno.int.cumu.post)#no differences between models
Anova(geno.int.cumu.post)#No interaction effect present, continue with simple genotype model
Anova(geno.cumu.post)
fixef(geno.cumu.post)
``